\section{How to build tfe (text file
editor)}\label{how-to-build-tfe-text-file-editor}

\subsection{How to compile and execute the text editor
`tfe'.}\label{how-to-compile-and-execute-the-text-editor-tfe.}

First, source files are in the
\href{https://github.com/ToshioCP/Gtk4-tutorial}{Gtk4-tutorila
repository}. How to download them is written at the end of the previous
section.

The following is the instruction of compilation and execution.

\begin{itemize}
\tightlist
\item
  You need meson and ninja.
\item
  If you have installed gtk4 from the source, you need to set
  environment variables to suit your installation.
\item
  Change your current directory to \passthrough{\lstinline!src/tfe5!}
  directory.
\item
  Type \passthrough{\lstinline!meson setup \_build!} for configuration.
\item
  Type \passthrough{\lstinline!ninja -C \_build!} for compilation. Then
  the application \passthrough{\lstinline!tfe!} is built under the
  \passthrough{\lstinline!\_build!} directory.
\item
  Type \passthrough{\lstinline!\_build/tfe!} to execute it.
\end{itemize}

Then the window appears. There are four buttons,
\passthrough{\lstinline!New!}, \passthrough{\lstinline!Open!},
\passthrough{\lstinline!Save!} and \passthrough{\lstinline!Close!}.

\begin{itemize}
\tightlist
\item
  Click on \passthrough{\lstinline!Open!} button, then a file chooser
  dialog appears. Choose a file in the list and click on
  \passthrough{\lstinline!Open!} button. Then the file is read and a new
  Notebook Page appears.
\item
  Edit the file and click on \passthrough{\lstinline!Save!} button, then
  the text is saved to the original file.
\item
  Click \passthrough{\lstinline!Close!}, then the Notebook Page
  disappears.
\item
  Click \passthrough{\lstinline!Close!} again, then the
  \passthrough{\lstinline!Untitled!} Notebook Page disappears and at the
  same time the application quits.
\end{itemize}

This is a very simple editor. It is a good practice for you to add more
features.

\subsection{Total number of lines, words and
characters}\label{total-number-of-lines-words-and-characters}

\begin{lstlisting}
$ LANG=C wc tfe5/meson.build tfe5/tfeapplication.c tfe5/tfe.gresource.xml tfe5/tfenotebook.c tfe5/tfenotebook.h tfetextview/tfetextview.c tfetextview/tfetextview.h tfe5/tfe.ui
   10    17   294 tfe5/meson.build
  110   334  3601 tfe5/tfeapplication.c
    6     9   153 tfe5/tfe.gresource.xml
  144   390  3668 tfe5/tfenotebook.c
   15    21   241 tfe5/tfenotebook.h
  235   821  8473 tfetextview/tfetextview.c
   32    54   624 tfetextview/tfetextview.h
   61   100  2073 tfe5/tfe.ui
  613  1746 19127 total
\end{lstlisting}
